Determinism, Complexity, and Predictability in Computer Performance

### 2014-03-10 methods.tex
editing
putting it in Phys Rev E Format
more pre formatting
removing windows
adding window footnote

\section{Experimental Methods and Time Series Explanations}\label{sec:methods}

% \begin{enumerate}
% \item \cmark  Experimental methods: (how we collect the %time series and what the times series areThis should be HPM %PAPI, which programs we model
%\item \cmark program descriptions and relevant citations
%\subitem \cmark description of \col
%\subitem \cmark description of \gcc
%\subitem \cmark description of \svd 
%\subitem \cmark description of svd regimes
% \end{enumerate}

@@

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.75\textwidth]{figs/colFullTS}
    \caption{The instructions executed per CPU clock cycle (IPC) during the execution of \col. Each point is the average IPC during a 100,000 instruction period.}
    \label{fig:col-ts}

@@

  \begin{figure}[t]
  \centering
    \includegraphics[width=0.75\textwidth]{figs/gccfullts}
    \caption{The instructions executed per CPU clock cycle (IPC) during the execution of \gcc. Each point is the average IPC during a 100,000 instruction period.}
    \label{fig:gcc-ts}
  \end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.75\textwidth]{figs/SVD1RegimesColored}
    \caption{The instructions executed per CPU clock cycle (IPC) during the execution of \svd. Each point is the average IPC during a 100,000 instruction period. The colors correspond to the six different regimes detected visually.}
    \label{fig:svd-ts-colored}
  \end{figure}

\svd brings up a very interesting point: as we observe a system the generating process can change in complexity over time. While it is clear in Figures \ref{fig:col-ts} and \ref{fig:gcc-ts} that a single system is being observed, it also appears \emph{visually} that the complexity of these two time series is consistent over time, i.e., while they are both complicated they don't appear to increase or decrease in complexity over the range of the program. \svd (seen in Figure \ref{fig:svd-ts-colored}) is different however, at least visually. Like Figures \ref{fig:col-ts} and \ref{fig:gcc-ts}, Figure \ref{fig:svd-ts-colored} is a single time series of the IPC of \svd, however as time progresses it is clear that something drastically changes in the underlying generating process and the structure of the time series  visually switches between ``regimes".

This is caused by the code of \svd moving between different subroutines as the SVD algorithm evolves, something not present in either \gcc or \col. We call these changes in \svd dynamics \emph{\svd regimes} and we have colored each of the six regimes differently in Figure \ref{fig:svd-ts-colored}. The advantage to splitting \svd into these regimes is to explore how complexity and predictability evolve over time for a system in drift. In Section \ref{sec:wpeRegime} we explore and validate the choice of these visually selected regime windows extending techniques from \cite{cao2004det}. Note, the purpose of this paper is \emph{not} to rigorously explore regime detection but to explore quantifying complexity of a time series. As such, making these regime splits serves to provide an additional 90 unique time series\footnote{Six Regimes with 15 individual runs each.} to explore. For notational convenience we refer to these time series as {\tt dgesdd$_i$} with $i \in \{1\dots6\}$ where $i$ corresponds to a regime of \svd, ordered from left to right.

>>>>

\section{Experimental Methods and Time Series Explanations}\label{sec:methods}
{\color{blue} EDITABLE}
% \begin{enumerate}
% \item \cmark  Experimental methods: (how we collect the %time series and what the times series areThis should be HPM %PAPI, which programs we model
%\item \cmark program descriptions and relevant citations
%\subitem \cmark description of \col
%\subitem \cmark description of \gcc
%\subitem \cmark description of \svd 
%\subitem \cmark description of svd regimes
% \end{enumerate}

@@

\begin{figure}[htbp]
  \centering
    \includegraphics[width=\columnwidth]{figs/colFullTS}
    \caption{The instructions executed per CPU clock cycle (IPC) during the execution of \col. Each point is the average IPC during a 100,000 instruction period.}
    \label{fig:col-ts}

@@

  \begin{figure}[t]
  \centering
    \includegraphics[width=\columnwidth]{figs/gccfullts}
    \caption{The instructions executed per CPU clock cycle (IPC) during the execution of \gcc. Each point is the average IPC during a 100,000 instruction period.}
    \label{fig:gcc-ts}
  \end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=\columnwidth]{figs/SVD1RegimesColored}
    \caption{The instructions executed per CPU clock cycle (IPC) during the execution of \svd. Each point is the average IPC during a 100,000 instruction period. The colors correspond to the six different regimes detected visually.}
    \label{fig:svd-ts-colored}
  \end{figure}

\svd brings up a very interesting point: as we observe a system the generating process can change in complexity over time. While it is clear in Figures \ref{fig:col-ts} and \ref{fig:gcc-ts} that a single system is being observed, it also appears \emph{visually} that the complexity of these two time series is consistent over time, i.e., while they are both complicated they don't appear to increase or decrease in complexity over the range of the program. \svd (seen in Figure \ref{fig:svd-ts-colored}) is different however, at least visually. Like Figures \ref{fig:col-ts} and \ref{fig:gcc-ts}, Figure \ref{fig:svd-ts-colored} is a single time series of the IPC of \svd, however as time progresses it is clear that something drastically changes in the underlying generating process and the structure of the time series  \emph{visually} switches between ``regimes".\footnote{It should be noted that a windowed version of the metric we introduce in Section~\ref{sec:meaComplex} has been used to detect siezures in brainwave data \cite{cao2004det}, which can be viewed as regime detection. When we used this method on \svd the regimes we chose visually had different entropies from regime to regime but but the regime boundaries seem fractal in nature. This is something we plan to investigate in future work. For the purposes of this paper the regime detection was done visually.}

This is caused by the code of \svd moving between different subroutines as the SVD algorithm evolves, something not present in either \gcc or \col. We call these changes in \svd dynamics \emph{\svd regimes} and we have colored each of the six regimes differently in Figure \ref{fig:svd-ts-colored}. The advantage to splitting \svd into these regimes is to explore how complexity and predictability evolve over time for a system in drift.  Note, the purpose of this paper is \emph{not} to rigorously explore regime detection but to explore quantifying complexity of a time series. As such, making these regime splits serves to provide an additional 90 unique time series\footnote{Six Regimes with 15 individual runs each.} to explore. For notational convenience we refer to these time series as {\tt dgesdd$_i$} with $i \in \{1\dots6\}$ where $i$ corresponds to a regime of \svd, ordered from left to right.

For regimes \cite{cao2004det}
